---
title: "SMU MSDS 6306 Case Study 1: Beers, Breweries, and Bitterness"
author: "Adil Siraj, Garrett Mozey, Dana Geislinger, & Luay Dajani"
date: "February 17, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Breweries

Study of popular Brewed Beers from American domestic and craft varieties; shining a light on what makes them similar and different.  The study analyses the Brews and styles from Lagers, Ales, to the sweet ciders in addition to their bitterness, alcohol content, and where their respective breweries.  From this study we look to find out trends and cultures depending on location and taste, and possibly their appeal to the consumers.

Goal is to analyze collected data, normalize and clean the data, and create analysis that will show an apperent relationship between the bitterness of the bear the alcoholic content. In addition show the bitterness and alcoholic content as per the location it was brewed to determine a trend in culture to that location.

Repository: https://github.com/danageis/SMU-MSDS6306-CaseStudy1

## Analysis 1: Number of Breweries per US State
In this analysis, raw brewery data is imported and analyzed to find the number of breweries in each state.

```{r num_breweries}
# Import breweries data and preview data structure.
breweries = read.csv("Raw Data/Breweries.csv", header=TRUE)
str(breweries)

# Entries in 'State' column start with a space.
# Clean data by removing leading white-space and verify new structure.
breweries$State = as.factor(trimws(breweries$State))
str(breweries)

# Print number of breweries in each state.
# Since each row is a brewery, the number of occurences of each state equals the number of breweries in that state.
breweries.by.state = tapply(breweries$State, breweries$State, length)
print(breweries.by.state)
```

The preceeding table lists the number of breweries located in each state. There are 558 breweries total in the data set, and the number of breweries in each state ranges from 1 to 47.

## Analysis 2: Merge of Beer & Breweries Data
In this analysis, raw beer data is imported and merged with the brewery data from the previous analysis. The first and last 6 observations of this merged data set are printed to verify the structure of the merged dataset.

```{r data_merge}
# Import beer data and preview data structure.
beers = read.csv("Raw Data/Beers.csv", header=TRUE)
str(beers)

# Change Brewery ID column name for beer data to match brewery data.
names(beers)[5] = "Brew_ID"

# Merge beer and brewery data by 'Brew_ID' and preview naming structure.
brew.data = merge(breweries, beers, by="Brew_ID", all=TRUE)
names(brew.data)

# Rename variables to be meaningful and verify new structure.
names(brew.data) = replace(names(brew.data), c(2, 5), c('Brew_Name', 'Beer_Name'))
str(brew.data)

# Print first 6 and last 6 observations of the merged data for verification.
head(brew.data, 6)
tail(brew.data, 6)
```

The beer and brewery datasets have been merged (by Brewery ID) successfully. To verify the structure of the merged data, the first and last 6 rows are printed above.

## Analysis 3: Missing Data (NA's in each column)

The analysis below examines the data to see which values are missing per observation or per row in order to determine how representative the data will be after all subsequent analysis is completed.

```{r data_missing}
# Create Vector Dataframe which indicates the missing data (NA's)  in each columns as TRUE or FALSE.
brew.data.missing = is.na(brew.data)

# Null out vector to be used in for loop
number_NAs = NULL

# Though not intened, creates vector of "Named ints" where the name of the row associated with each int value
for (i in colnames(brew.data.missing))
{
  tempvec = brew.data.missing[,i]
  number_NAs[i] = length(tempvec[tempvec != FALSE])
}

print(number_NAs)
```

From a total of 2,004 total different brews. The results above indicate that out of the 2,004 total, 62 of the beers will not have an indication of its Alcohol By Volume (ABV) and 1,005 beers will not indicate their International Bitterness Unit (IBU).  This means that trying to make assumptions on ABV and IBU will only be valid for around half the data collected.

## Analysis 4: Compute median alcohol content (ABV) and international bitterness unit (IBU) per each state.

The analysis below will compute the mid-range (median) of the alcohol content (ABV) and international bitterness unit (IBU) per state where that data is available to determine if there is a specific trend or culture associated with the breweries in a specific state.

```{r data_abv_ibu, fig.width=10, fig.height=9}
# Use "tapply" to find the the median IBU per state
ABV_medians = tapply(brew.data$ABV   # Take the IBU column (numeric)
                     , brew.data$State # Use the State column as the index (as factor) to associate IBU per State
                     , function (x) {median(x, na.rm = TRUE)} #Run created function finding the median per State removing NULLs
)

#Print out the results
cat("The Median Alcohol Content (ABV) per state:\n") #using cat to preserve the \n
print(ABV_medians)

#barplot of the ABV medians
par(las = 2) #make the x axis name horizontal with the horizontal view
par(mar=c(5,10,4,2)) # increase y-axis margin (c(bottom, left, top, right))
barplot(ABV_medians
        , main="Median Alcohol Content (ABV) per State"
        , xlab="Median ABV"
        , ylab="States"
        , cex.names=0.7 #Font Size for the State names to squeeze them all in
        , col=c("red","green","blue","orange","brown","purple","yellow")
        , horiz=TRUE
     )

# Use "tapply" to find the the median IBU per state
IBU_medians = tapply(brew.data$IBU   # Take the IBU column (numeric)
       , brew.data$State # Use the State column as the index (as factor) to associate IBU per State
       , function (x) {median(x, na.rm = TRUE)} #Run created function finding the median per State removing NULLs
)

#Print out the results
cat("The International Bitterness Unit (IBU) per each state:\n") #using cat to preserve the \n
print(IBU_medians)

#barplot of the ABV medians
par(las = 2) #make the y axis name horizontal with the horizontal view
par(mar=c(5,10,4,2)) # increase y-axis margin (c(bottom, left, top, right))
barplot(IBU_medians
        , main="Median International Bitterness Unit (IBU)  per State"
        , xlab="Median IBU"
        , ylab="States"
        , cex.names=0.7 #Font Size for the State names to squeeze them all in
        , col=c("red","green","blue","orange","brown","purple","yellow")
        , horiz=TRUE
)

```

As per the bar plots above, the states with the highest alcohol content are DC (Washington DC) then Kentucky, while the states brewing bear with a lesser alcohol content would be Utah then New Jersey.

Likewise, the states making the most bitter bear are Maine followed by West Virginia, while the states brewing the least bitter bear are Wisconsin then Kansas, with South Dakota not having any bitterness data for their brews.

## Analysis 5: Show which state has the maximum alcoholic (ABV) beer and well as the most bitter.

```{r data_state_statistics}
### to get the Maximum Alcoholic beer per state
ABV=brew.data[c('State','ABV')] # Obtain data for alcoholic contents per state
ABV[which.max(brew.data$ABV),]['State']  # get the maximum of Alcoholic content out of all states

IBU=brew.data[c('State','IBU')] # Obtain data for bitterness in beer per state
IBU[which.max(brew.data$IBU),]['State'] # get the most bitter beer out of all states

```
Colorado has beer with the maximum alcoholic contents of 0.126 out of all states
Oregon has the most bitter beer taste out of all states with international bitterness unit of 138

## Analysis 6: Summary of statistics for the alcohol content (ABV).

```{r data_abv_stats}
#Summary of Alcoholic contents in beer
summary(ABV$ABV)
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#0.00100 0.05000 0.05600 0.05977 0.06700 0.12800      62 

```
As per the summary, Maximum Alcoholic content in beer is 0.128 which is in Colorado and mimimum is in California which is .001

## Analysis 7: Prove relationship between the bitterness of the beer and its alcoholic content.

The plot analysis below depicts the relationship between the bitterness of the beer
and its alcoholic content using a scatter plot with a smoothed line that is meant
to easily demonstrate any apparent data pattern.

```{r ibu_vs_abv}
# Invoking library to use ggplot2
library(ggplot2)

# Assigning 'IBU' and 'ABV' columns from brew data frame to 'IBU_ABV' variable
IBU_ABV = brew.data[c('IBU','ABV')]

# Setting up basic plot 
basic_plot = ggplot(IBU_ABV, # Assign data using 'IBU_ABV' variable 
                  aes(IBU,ABV)) # Construct aesthetic mappings for two data columns

# Add scatter points to basic plot along with axis labels and a plot title  
scatter_with_labels = basic_plot + 
                      geom_point(na.rm = TRUE) +  # scatter points added 
                      labs(x = "IBU", 
                           y = "ABV", 
                           title= "Linear Regression")

# Add regression line smoothing to aid in seeing patterns
scatter_smoothed = scatter_with_labels + 
                    stat_smooth(method="auto"  # line smoothing call
                                ,na.rm = TRUE) # silently remove missing values without warning 

# Display the scatter plot
scatter_smoothed 

```
As shown above with the smoothed line, the general pattern indicates that as alcoholic volume (ABV) increases, the bitterness also increases. Once the alcoholic volume (ABV) approaches about 0.08 units and the bitterness (IBU) reaches 100 units, there are far less beer samples. Most of the beer samples tend to be relatively less bitter and lower in alcoholic volume.


## Conclusion
In this study, data from United States breweries and the beers they produce were merged into a single dataset and used to investigate the distribution of breweries across the country.

## Environment Notes

```{r environment_notes, echo=FALSE}
sessionInfo()

```
