---
title: "SMU MSDS 6306 Case Study 1: Beers, Breweries, and Bitterness"
author: "Adil Siraj Garrett Mozey Dana Geislinger Luay Dajani"
date: "February 17, 2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Add all initialization and Library files here

```

# Breweries

Study of popular Brewed Beers from American domestic and craft varieties; shining a light on what makes them similar and different.  The study analyses the Brews and styles from Lagers, Ales, to the sweet ciders in addition to their bitterness, alcohol content, and where their respective breweries.  From this study we look to find out trends and cultures depending on location and taste, and possibly their appeal to the consumers.

Goal is to analyze collected data, normalize and clean the data, and create analysis that will show an apperent relationship between the bitterness of the bear the alcoholic content. In addition show the bitterness and alcoholic content as per the location it was brewed to determine a trend in culture to that location.

## Analysis 1: Number of Breweries per US State

```{r num_breweries}
# Import breweries data and preview data structure.
breweries = read.csv("Raw Data/Breweries.csv", header=TRUE)
str(breweries)

# Entries in 'State' column start with a space.
# Clean data by removing leading white-space and verify new structure.
breweries$State = as.factor(trimws(breweries$State))
str(breweries)

# Print number of breweries in each state.
# Since each row is a brewery, the number of occurences of each state equals the number of breweries in that state.
breweries.by.state = tapply(breweries$State, breweries$State, length)
print(breweries.by.state)

```

## Analysis 2: Merge of Beer & Breweries Data

```{r data_merge}
# Import beer data and preview data structure.
beers = read.csv("Raw Data/Beers.csv", header=TRUE)
str(beers)

# Change Brewery ID column name for beer data to match brewery data.
names(beers)[5] = "Brew_ID"

# Merge beer and brewery data by 'Brew_ID' and preview naming structure.
brew.data = merge(breweries, beers, by="Brew_ID", all=TRUE)
names(brew.data)

# Rename variables to be meaningful and verify new structure.
names(brew.data) = replace(names(brew.data), c(2, 5), c('Brew_Name', 'Beer_Name'))
str(brew.data)

# Print first 6 and last 6 observations of the merged data for verification.
head(brew.data, 6)
tail(brew.data, 6)

```

## Analysis 3: Missing Data (NA's in each column)

```{r data_missing}
#

```

## Analysis 4: Compute median alcohol content (ABV) and international bitterness unit (IBU) per each state.

```{r data_abv_ibu}
#

```

## Analysis 5: Show which state has the maximum alcoholic (ABV) beer and well as the most bitter.

```{r data_state_statistics}
#

```

## Analysis 6: Summary of statistics for the alcohol content (ABV).

```{r data_abv_stats}
#

```

## Analysis 7: Prove relationship between the bitterness of the beer nad its alcoholic content.

```{r ibu_vs_abv}
#

```